\bf Approximating Likelihood Ratios with Calibrated Discriminative Classifiers

### 2016-02-08 jasa/paper-jasa.tex

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xfrac}
\usepackage{hyperref}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

\subsection{Mixture models}

@@

\glnote{todo}
\glnote{+ diagnostics}

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xfrac}
\usepackage{hyperref}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{subcaption}

@@

\subsection{Mixture models}
\label{sec:mixture}

@@

As a simple but illustrative example, let us first consider the approximation of
the likelihood log-ratio $\log \left( r(\mathbf{x};\theta_0,\theta_1) \right)$ between the mixtures
$p(\mathbf{x}|\theta_0)$ and $p(\mathbf{x}|\theta_1)$ illustrated in
Figure~\ref{fig:1} and defined as
\begin{align}
p(\mathbf{x}|\theta_0) &= (\frac{1}{2} - g) p_{c_0}(\mathbf{x}) + (\frac{1}{2} - g) p_{c_1}(\mathbf{x}) + g p_{c_2}(\mathbf{x}), \\
p(\mathbf{x}|\theta_1) &= \frac{1}{2} p_{c_0}(\mathbf{x}) + \frac{1}{2} p_{c_1}(\mathbf{x}),
\end{align}
where $p_{c_0} := {\cal N}(\mu=-2, \sigma^2=0.5625)$, $p_{c_1} := {\cal N}(\mu=0, \sigma^2=4)$,
$p_{c_2} := {\cal N}(\mu=1, \sigma^2=0.25)$ and where $g$ is a parameter of $p(\mathbf{x}|\theta_0)$
set to $0.05$.

Using the method outlined in Section~\ref{sec:approx},
figures~\ref{fig:2}, \ref{fig:3} and \ref{fig:4} show the approximated
ratio when using respectively a linear model, a neural network or a random
forest. The blue curve corresponds to the true target log-ratio, the green curve is
the direct approximation $\log \left( \frac{\hat s(\mathbf{x})}{1 - \hat s(\mathbf{x})} \right)$
without calibration, while the red curve is the approximation $\log \frac{\hat p(\hat
s(\mathbf{x}) | \theta_0)}{\hat p(\hat s(\mathbf{x}) | \theta_1)}$ as calibrated
using kernel density estimation for a fixed bandwidth set to $0.2$. Finally, the
cyan curve shows the approximated ratio when decomposing the mixtures, as
described in Section~\ref{sec:mixture}.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig1.pdf}
        \caption{$p(\mathbf{x}|\theta_0)$ and $p(\mathbf{x}|\theta_1)$}
        \label{fig:1}
    \end{subfigure}
    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig2-a.pdf}
        \caption{Linear model}
        \label{fig:2}
    \end{subfigure}

    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig2-b.pdf}
        \caption{Neural network}
        \label{fig:3}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/fig2-c.pdf}
        \caption{Random forest}
        \label{fig:4}
    \end{subfigure}
    \caption{Approximation of the log-likelihood ratio $\log r(\mathbf{x};\theta_0,\theta_1) = \log \frac{p(\mathbf{x}|\theta_0)}{p(\mathbf{x}|\theta_1)}$ with calibrated classifiers. \glnote{I need to regenerate these figures using cross-validation for training and calibrating!}}\label{fig:animals}

\end{figure}

\glnote{Diagnostics.}

\glnote{Extend to illustrate a generalized likelihood ratio test with
parameterized classifier / calibration by assuming that the true value of $g$ is unknown.}
